---
title: "First pass at Maggie's AIM vs. Snapshot analysis"
output: html_notebook
---

Overview: First pass at Maggie's data analysis

set things up

```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(lubridate)
```

Pull in data
```{r}
df<-read.csv(here("Output", "PostZooniverse", "AssignedSpecies_prebait_incomplete.csv"))
```

Now pull in camera data
```{r}
ct<-read.csv(here("Data", "PostZooniverse","cameratrapdata2021.csv"))
```

Let's summarize the dataset and look for possible errors

```{r}
summary(df)
```
Let's convert Treatment to a factor variable and DTO to a date time object.

```{r}
#factor first
df$Treatment<-as.factor(df$Treatment)
df$DTO<-mdy_hm(df$DTO)
```

Treatment appears to have 3 levels but should only have 2. Let's fix that.

```{r}
unique(df$Treatment)
levels(df$Treatment)
```
Let's change all the ones with "Snap" to "Snapshot"

First, we have to find all the rows in which Treatment == Snap

```{r}
fix<-which(df$Treatment == "Snap")
df$Treatment[fix]<-"Snapshot"
df$Treatment<-factor(df$Treatment)
levels(df$Treatment)
```



Let's look at how many species were detected

```{r}
num_species<-length(unique(df$choice))
species<-unique(df$choice)
species
```
Now let's plot the number of each species as a function of treatment

```{r}
Species_summary<-df %>% group_by(choice, Treatment) %>% summarise(
  num_obs = length(choice)

)
```
Now make a plot of the number of observations of each species as a function of treatment

```{r}
ggplot(Species_summary, aes(choice, num_obs, fill = Treatment))+
  geom_col(position = "dodge")+
  labs(y = "Number of observations")+
  coord_flip()
```
Now it would be nice to be able to extract the camera and sd card number from DF so we can join to the ct data. Let's use str_sub to do so.

First, use `str_sub` to isolate camera number and sd card number in df

```{r}
df<-df %>% mutate(
  Cam = str_sub(Img1,3,4),
  SD = str_sub(Img1,9, 10),
  Cam_SD = paste0(Cam,"-", SD)
)
```
Now go to ct and do similar

```{r}
ct<-ct %>% mutate(
  Cam = Camera_number,
  SD = SD_num,
  Cam_SD = paste0(Cam,"-", SD)
)
```

Now we can join the two data frames together!
```{r}
mydata<-left_join(df,ct, by = "Cam_SD")
```
Now calculate trap nights; first convert character to date
```{r}
mydata$start_date<-dmy(mydata$start_date)
mydata$retrieval_date<-dmy(mydata$retrieval_date)
```
Now calculate trap nights

```{r}
mydata<-mydata %>% mutate(
  Trap_nights = retrieval_date-start_date
)
```

Now we need to calculate observations per trap night

```{r}
Species_summary<-mydata %>% group_by(choice, Cam_SD,Treatment) %>% summarise(
  num_obs = length(choice),
  trap_nights = max(Trap_nights)

)
```
Now calculate observation per 100 trap nights
```{r}
Species_summary<-Species_summary %>% mutate(
  obs_per_nt = num_obs/as.numeric(trap_nights),
  obs_per_100 = obs_per_nt*100
)
```

In order to plot, we need avg # of observation per 100 trap nights per species, not grouped by camera

```{r}
Obs_summary<-Species_summary %>% group_by(choice, Treatment) %>% summarise(
  avg_obs_100 = mean(obs_per_100,na.rm = T),
  sd_obs_100 = sd(obs_per_100,na.rm =T)
)
```

Let's plot!

```{r}
ggplot(Obs_summary, aes(choice, avg_obs_100, fill = Treatment))+
  geom_col(position = "dodge")+
  coord_flip()+
  labs(x = "Species", y = "Avg. observations per 100 nights")
```
Let's do a statistical test (ANOVA) comparing the average number of observations between treatments, regardless of species.
```{r}
Treatment<-Species_summary %>% group_by(Cam_SD, Treatment) %>% summarise(
  Obs_per_100 = sum(obs_per_100, na.rm = T)
)
```
Now let's run an anova

```{r}
my_model<-glm(Obs_per_100~Treatment, data = Treatment)
summary(my_model)
anova(my_model)
```
Now let's see if there is a difference in the number of species detected between the two methods

```{r}
NumSpecies<-Species_summary %>% group_by(Cam_SD,Treatment) %>% summarise(
  NumSpecies = length(choice)
)
```

Now run a glm to see if there is a statistical difference in the # of species per treatment

```{r}
num_model<-glm(NumSpecies~Treatment, data = NumSpecies)
summary(num_model)
anova(num_model)
```


