---
title: "R Notebook"
output: html_notebook
---
Main analysis of Maggie's data

Begin by setting things up:
```{r}
rm(list = ls())
library(tidyverse)
library(lubridate)
library(here)
```
1. Is there a difference in how many images were misfires or in total images taken between treatments, phases?

pull in is_there_a_species data
```{r}
DF<-read.csv(here("Data", "Maggie_clean_is_there_species.csv"))
```
Get a summary by camera, phase, treatment of number of images of each type (misfire, animal present)

```{r}
Camera_summary<-DF %>% group_by(CamSD, Phase, Treatment,  value) %>% summarise(
 num_responses = length(value)
)
```
Now get total number of events so that I can calculate percent of images that are misfires
```{r}
totals<-DF %>% group_by(CamSD, Phase, Treatment) %>% summarise(
  num_events = length(value)
)
```
Join num_events to Cam summary
```{r}
Camera_summary<-left_join(Camera_summary, 
                          totals %>% dplyr::select(num_events),
                          by = "CamSD")
Camera_summary<-Camera_summary %>% select(-Phase.y)
Camera_summary<-Camera_summary %>% mutate(
  Percent = round(100*(num_responses/num_events), 2)
)
```
now filter
```{r}
Percent_misfires<-Camera_summary %>% filter(value == "I don't think so")
```

Now see if there is a difference in the % misfires b/w phases, treatments. Use glm with binomial family since dealing with percent data.

Need a new binomial column based on value

```{r}
Camera_summary$binom<-NA
fix<-which(Camera_summary$value == "Yes")
Camera_summary$binom[fix]<-1
fix<-which(is.na(Camera_summary$binom))
Camera_summary$binom[fix]<-0
```

Now run binomial glm (logistic regression...)
```{r}
per_mis<-glm(binom~Phase.x/Treatment, data = Camera_summary, family = binomial)
summary(per_mis)
anova(per_mis)
```

Plot
```{r}
cam_fire_summary<-Camera_summary %>% arrange(value, Treatment, Phase.x) %>% 
      group_by(Phase.x, Treatment, value) %>% summarise(
        tot_responses = sum(num_responses, na.rm = T),
        avg_responses = mean(num_responses, na.rm = T), 
        sd_responses = sd(num_responses, na.rm = T)
          )
```

And make the graph
```{r}
ggplot(cam_fire_summary, aes(Treatment, avg_responses, color = value, group = value))+
  geom_point()+
  geom_line()+
  facet_wrap(~Phase.x)
```

