---
title: "Data cleaning"
output: html_notebook
---

Cleaning up some of Maggie's data for general use.

Set things up.

```{r}
rm(list = ls())
library(tidyverse)
library(lubridate)
library(here)
```

## 1. Camera information
Pull in data

```{r}
cams<-read.csv(here("Data", "cameratrapdata2021.csv"))
summary(cams)
```
Set dates
```{r}
cams$start_date<-dmy(cams$start_date)
cams$retrieval_date<-dmy(cams$retrieval_date)
summary(cams$start_date) #looks good
summary(cams$retrieval_date)
```
Max of 2022-11-16 is incorrect, as that date hasn't happened yet. That date should be corrected to 2021-11-16

```{r}
fix<-which(cams$retrieval_date == "2022-11-16")
cams$retrieval_date[fix]<-"2021-11-16"
summary(cams$retrieval_date)
```
Now looks good. 

Now check treatment, phase for factor levels
```{r}
unique(cams$phase)
unique(cams$protocol)
```
Both look good.

Now check forest names, habitat, setting
```{r}
unique(cams$forest)
unique(cams$habitat)
unique(cams$setting)
```
Now make snowstake distance into a numeric rather than character variable by removing the "m".

```{r}
cams$snow_stake_dist_m<-str_remove(cams$snowstake_dist, "m")
cams<-select(cams, -snowstake_dist)
cams<-cams %>% relocate(snow_stake_dist_m,.after = trig_delay_sec)
```
Looks good. 

Now check lat/long and change column names

Now save
```{r}
fixem<-colnames(cams)
fixem[20]<-"Lat"
fixem[21]<-"Long"
colnames(cams)<-fixem
```
Now make all longitudes negative
```{r}
fix<-which(cams$Long > 1)
cams$Long[fix]<-cams$Long[fix]*-1
```
and now check that there is only one lat/long per camera location (cams did not move but changed sd cards so can group by cams to see)

```{r}
cam_locations<-cams %>% group_by(Camera_number) %>% summarise(
  num_Lats=length(unique(Lat)),
  num_Longs = length(unique(Long))
)
```
Looks good. Now save!

```{r}
write.csv(cams, here("Data", "maggie_clean_cam_data.csv"))
```




## 2. Is there a species data
Read in the data
```{r}
is_there<-read.csv(here("Output", "PostZooniverse","Flattened", "flat_is_there_animal-2022-05-02.csv"))
```
Remove X column
```{r}
is_there<-select(is_there, -X)
```
Check phase and treatment

```{r}
unique(is_there$Phase)
```
Need to combine batch 1 and batch 2 of phase 1 into just phase 1.

```{r}
fix<-which(is_there$Phase == "phase 1 batch 1"|is_there$Phase == "phase 1 batch 2" )

is_there$Phase[fix]<-"phase 1"
unique(is_there$Phase)
```
Now look at treatment
```{r}
unique(is_there$Treatment)
```
Change "Snap" into "Snapshot"
```{r}
fix<-which(is_there$Treatment == "Snap")
is_there$Treatment[fix]<-"Snapshot"
unique(is_there$Treatment)
```
Check forest names
```{r}
unique(is_there$Forest)
```
Need to combine Whip into Whippoorwill
```{r}
fix<-which(is_there$Forest == "Whip")
is_there$Forest[fix]<-"Whippoorwill"
unique(is_there$Forest)
```
Now check dates and make into date field
```{r}
is_there$DTO<-mdy_hm(is_there$DTO)
summary(is_there$DTO)
```
There are observations from 2017 which are likely not correct. Check them out.
```{r}

ggplot(is_there, aes(year(DTO)))+
         geom_histogram()
```
There are a few images from 2019 and 2017 that must be camera setting errors. Let's check them out. 
```{r}
fix<-which(year(is_there$DTO) < "2020")
check<-is_there %>% slice(fix)
fix<-which(year(is_there$DTO) == "2017")
is_there$DTO[c(1168,1466,1830)]<-"2021-10-17 2:50:00"
is_there$DTO[c(2207,2410,2443)]<-"2021-10-17 2:51:00"
fix<-which(year(is_there$DTO) == "2019")
is_there$DTO[fix]<-"2021-10-24 4:39:00"
summary(is_there$DTO)
```

Looks good. Now see if there are any cameras for which there is > 1 lat/long listed

```{r}
cam_locations<-is_there %>% group_by(CamSD) %>% summarise(
 num_lat = length(unique(Lat)),
 num_long = length(unique(Long))
)
```

There is an issue with camera 49 sd 60

```{r}
c49_sd60<-is_there %>% filter(CamSD == "C049_SD060") %>% summarise(
  camLats = unique(Lat), 
  camLongs = unique(Long)
)
```
Numbers are quite different - need to sort out. Correct numbers are Lat = 44.39964792, Long = 75.12164175

```{r}
fix<-which(is_there$CamSD == "C049_SD060" & is_there$Lat > 44.4)
is_there$Lat[fix]<- 44.39964792
is_there$Long[fix]<- -75.12164175
```

Ok! All fixed. Now save

```{r}
write.csv(is_there, here("Data", "Maggie_clean_is_there_species.csv"))
```


## 3. Species data
Pull in data
```{r}
species<-read.csv(here("Output", "PostZooniverse", "Maggie_final_species.csv"))
```

Get rid of X.1 and X columns
```{r}
species<-species[, 3:14]
```

how are species choices defined?
```{r}
unique(species$choice)
```

Looks good.
Add a column for taxon
```{r}
species$taxon <-NA
choices<-unique(species$choice)
birds<-c("BLUEJAY", "OTHERBIRD", "GROUSERUFFED", "TURKEY", "OWL")
fix<-which(species$choice %in% birds)
species$taxon[fix]<-"bird"
mammals<-(which(!(species$choice %in% birds)))
species$taxon[mammals]<-"mammal"
```

Add size and diet according to Maggie's classifications

```{r}
size<-read.csv(here("Data", "size_diet.csv"))
ourchoices<-unique(species$choice)
which(size$SPECIES %in% ourchoices)
size$SPECIES[17]<- "unknown"
```

Now join to get a column of size and diet based on species

```{r}
species<-left_join(species,
                   size %>% dplyr::select(DIET, AVG_SIZE_kg), 
                   by = c("choice" = "SPECIES")
                   )
```

Try chaning column name in size to match choice
```{r}
fixem<-colnames(size)
fixem[1]<-"choice"
colnames(size)<-fixem
```
Now try the join
```{r}
species<-left_join(species,
                   size %>% dplyr::select(DIET, AVG_SIZE_kg),
                   by = "choice")
```

This should be working and isn't so I'm moving on to a different way.

```{r}
size<-size %>% select(choice, DIET, AVG_SIZE_kg)
species<-left_join(species, size)
```

check phases
```{r}
unique(species$Phase)
```

Need to merge phase 1 batches together
```{r}
fix<-which(species$Phase == "phase 1 batch 1" | species$Phase == "phase 1 batch 2")
species$Phase[fix]<-"Phase 1"
unique(species$Phase)
```

Check treatments
```{r}
unique(species$Treatment)
```

Need to combine Snapshot and Snap
```{r}
fix<-which(species$Treatment == "Snap")
species$Treatment[fix]<- "Snapshot"
unique(species$Treatment)
```

Now fix DTO
```{r}
species$DTO<-mdy_hm(species$DTO)
summary(species$DTO)
```

As with before, I have dates in 2017 and 2019 that need fixing.

```{r}
fix<-which(year(species$DTO) == "2017")
species$DTO[43]<-"2021-10-17 2:50:00"
species$DTO[44]<-"2021-10-17 2:51:00"

fix<-which(year(species$DTO)== "2019")
species$DTO[fix]<-"2021-10-24 4:39:00"
```


Ok. Now deal with lat/long
Start by adding camera/sd column to species
```{r}
species$SD<-str_sub(species$Img1, 6,10)
species$Cam_SD<-paste0(species$CamNum, "_", species$SD)
```

```{r}
place_sum<-species %>% group_by(Cam_SD) %>% summarise(
  num_lats = length(unique(Lat)),
  num_longs = length(unique(Long))
)
```

Again need to fix cam 49 sd 60

```{r}
fix<-which(species$Cam_SD == "C049_SD060" & species$Lat > 44.4)
species$Lat[fix]<- 44.39964792
species$Long[fix]<- -75.12164175
```
Now write cleaned data set
```{r}
write.csv(species, here("Data", "maggie_clean_species.csv"))
```


