---
title: "Analyze Is there an animal?"
output: html_notebook
---
## Introduction
R code to analyze Maggie's data from the "Is there an animal" workflow to find out how many misfires, etc.

Set things up:

```{r, message=FALSE, echo=FALSE}
rm(list = ls())
library(here)
library(tidyverse)
library(tidyjson)#may get error; if so, load from github
library(jsonlite)
library(magrittr) #allows piping beyond tidyverse
library(dplyr)
library(stringr)
library(tidyr)
library(lubridate)
library(rstudioapi)

myFunctions<-list.files(here("Code", "Functions"), pattern = "R") 
#points to files that need to be sourced

sapply(here("Code", "Functions", myFunctions), source) #sources all of the functions in the Functions folder.
```
###Choose the flat file to work with.
We have multiple classification files. The choose_file_isol asks the user for input so that they can select the classification file to work with. The classification file should be on the t://drive in Bart_s04/Research_group/NoCoWild/Data/classification_files.

Try with the `choose_my_classification_file` function

```{r}
filepath<-here("Output","PostZooniverse", "Flattened")
DF<-choose_my_classification_file(filepath)
data<-DF[[2]]
filename<-"flat_is_there_animal"
```
### Combine Phase 1 batches into one
There are two categories for phase 1 that should be combined into a new category just called "phase 1".

```{r}
unique(data$Phase)
data<-data %>% mutate(
  RealPhase = Phase
)

fix<-which(data$Phase == "phase 1 batch 1" | data$Phase == "phase 1 batch 2")

data$RealPhase[fix]<- "phase 1"

unique(data$RealPhase)
```
Now fix the levels of treatment so there are only 2

```{r}
unique(data$Treatment)
fix<-which(data$Treatment == "Snap")
data$Treatment[fix]<- "Snapshot"
unique(data$Treatment)
```

## Analysis Questions

### How many camera fires
How many times did the camera fire, in total and broken down by treatment and phase

```{r}
Fires<-data %>% group_by(RealPhase, Treatment, value) %>% summarise(
  NumFires = length(task_label)
  
)
```
Add a column that combines Phase and Treatment
```{r}
Fires<-Fires %>% mutate(
  Phase_treat = paste0(RealPhase, "-", Treatment)
)
```

And plot
```{r}
ggplot(Fires, aes(Phase_treat,NumFires, fill = value ))+
  geom_col(position = "dodge")+
  
  coord_flip()
```



