---
title: "Species analysis"
output: html_notebook
---

## Introduction
This is the .Rmd file for doing the final species analysis for Maggie's Aim_v_Snap project. It uses the file called "Maggie_final_species.csv" located in the project Output folder as the data.

## Initial set up and tidy up
Set things up and pull in data
```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(lubridate)

DF<-read.csv(here("Output", "PostZooniverse", "Maggie_final_species.csv"))
```

Let's take a look at the data
```{r}
summary(DF)
```
Now let's clean things up

```{r}
DF<-DF[,3:14] #get rid of unwanted columns
unique(DF$Treatment)
```
Fix Snap to Snapshot

```{r}
fix<-which(DF$Treatment == "Snap")
DF$Treatment[fix]<-"Snapshot"
unique(DF$Treatment)
```
Now check species
```{r}
unique(DF$choice)
```
Looks good.

Now check Phase
```{r}
unique(DF$Phase)
```
Combine all the phase 1 types into one category

```{r}
fix<-which(DF$Phase == "phase 1 batch 1"|DF$Phase == "phase 1 batch 2")
DF$Phase[fix]<-"experiment"
unique(DF$Phase)
```
Now make a column that combines phase and treatment
```{r}
DF<-DF %>% mutate(
  Phase_Treat = paste0(Treatment,"-",Phase)
)
```

## Descriptive statistics

How many species are there by treatment and phase?

```{r}
Species<-DF %>% group_by(Treatment, Phase) %>% summarise(
  NumSpecies = length(unique(choice))
)
```
Let's take a look by camera as well:
```{r}
Species<-DF %>% group_by(Treatment, Phase, CamNum) %>% summarise(
  NumSpecies = length(unique(choice))
)
```

## Species - area curves

It would be nice to know the cumulative number of species as you increase camera trap nights. To get this, we need to join to Maggie's table about cameras.
```{r}
cams<-read.csv(here("Data", "cameratrapdata2021.csv"))
```
Now get a cam_sd column in DF to join by

```{r}
DF<-DF %>% mutate(
  Cam_SD = str_sub(Img1, 1,10)
)
```

Now join
```{r}
DF<-left_join(DF, cams, by = "Cam_SD")
```
Now let's calculate trap nights. Need to make some date vars to do so.

```{r}
DF$DTO<-mdy_hm(DF$DTO)
DF$start_date<-dmy(DF$start_date)
DF$retrieval_date<-dmy(DF$retrieval_date)
```

Now that we have those in as dates, let's get a summary of them:

```{r}
summary(DF$DTO)
```
Uh-oh! Some of the camera dates are incorrect, as we did not have cameras set in 2017. Let's take a look and try to fix that.

```{r}
bad_dates<-DF %>% filter(year(DTO) < 2021)
```
These are all human camera trappers. I'll try to fix, but can basically ignore if it doesn't work very well.

```{r}
good_dates<-ymd_hms(c("2021-10-17 14:50:00", "2021-10-17 14:51:00", "2021-10-04 16:23:00"))
subs<-bad_dates$subject_ids
fix<-which(DF$subject_ids %in% subs)
DF$DTO[fix]<-good_dates
summary(DF$DTO)
```
Now let's look at start and end dates
```{r}
summary(DF$start_date)
summary(DF$retrieval_date)
```
I can see that there is a retrieval date that is set for november 2022, a date that hasn't happened yet!  Need to fix that

```{r}
funky<-DF %>% filter(retrieval_date == "2022-11-16")
fixem<-funky$subject_ids
fix<-which(DF$subject_ids %in% fixem)
DF$retrieval_date[fix]<-ymd("2021-11-16")
summary(DF$retrieval_date)
```
Good!



Now write some code to calculate the cumulative number of species over days for each camera.

Begin by arranging DF by Cam_SD and then start_date
```{r}
DF<-DF %>% arrange(Cam_SD,start_date)
```

Need a function that will:
1. Grab the data for a single camera
2. Assign each date to a day number for counting the number of species on each day
3. Create a data frame to add data for each camera with columns: Cam_SD, phase_treatment, cumulative_day, cum_num_species
4. Loop through each camera and do the assigning.

Let's call it "species_cam_nights"

```{r}
species_cam_nights<-function(df){
  df<-df %>% arrange(Cam_SD, DTO)
  num_cams<-length(unique(df$Cam_SD))
  cams<-unique(df$Cam_SD)
  df$day<-NA
  
  #loop to assign day
  for(i in 1:num_cams){
    data<-df %>% filter(Cam_SD == cams[i])#grab the data
    num_rows<-length(data$Cam_SD)
    data$day[1]<-1 #assign first day as 1
      for(z in 2:num_rows){
        data$day[z]<-as.integer(round(difftime(data$DTO[z], data$DTO[z-1], units = "days")))
      }
    #now replace those subject ids in df
    sub_ids<-unique(data$subject_ids)
    swap<-which(df$subject_ids %in% sub_ids)
    df[swap,]<-data
  }
  df<-df %>% relocate(day, .after = DTO)
  return(df)
}
```

This function is working but I need to fix some dates on some of the cameras - the DTO is incorrect on some of the events.