---
title: "Species analysis"
output: html_notebook
---

## Introduction
This is the .Rmd file for doing the final species analysis for Maggie's Aim_v_Snap project. It uses the file called "Maggie_final_species.csv" located in the project Output folder as the data.

## Initial set up and tidy up
Set things up and pull in data
```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(lubridate)

DF<-read.csv(here("Output", "PostZooniverse", "Maggie_final_species.csv"))
```

Let's take a look at the data
```{r}
summary(DF)
```
Now let's clean things up

```{r}
DF<-DF[,3:14] #get rid of unwanted columns
unique(DF$Treatment)
```
Fix Snap to Snapshot

```{r}
fix<-which(DF$Treatment == "Snap")
DF$Treatment[fix]<-"Snapshot"
unique(DF$Treatment)
```
Now check species
```{r}
unique(DF$choice)
```
Looks good.

Now check Phase
```{r}
unique(DF$Phase)
```
Combine all the phase 1 types into one category

```{r}
fix<-which(DF$Phase == "phase 1 batch 1"|DF$Phase == "phase 1 batch 2")
DF$Phase[fix]<-"experiment"
unique(DF$Phase)
```
Now make a column that combines phase and treatment
```{r}
DF<-DF %>% mutate(
  Phase_Treat = paste0(Treatment,"-",Phase)
)
```

## Species - area curves

It would be nice to know the cumulative number of species as you increase camera trap nights. To get this, we need to join to Maggie's table about cameras.
```{r}
cams<-read.csv(here("Data", "cameratrapdata2021.csv"))
```
Check cams for issues

```{r}
summary(cams)
```
Make some dates
```{r}
cams$start_date<-dmy(cams$start_date)
cams$retrieval_date<-dmy(cams$retrieval_date)
summary(c(cams$start_date, cams$retrieval_date))
```
I can see that there is a retrieval date that is set for november 2022, a date that hasn't happened yet!  Need to fix that

```{r}
funky<-which(cams$retrieval_date == "2022-11-16")
cams$retrieval_date[funky]<-ymd("2021-11-16")
summary(cams$retrieval_date)
```


Now get a cam_sd column in DF to join by

```{r}
DF<-DF %>% mutate(
  Cam_SD = str_sub(Img1, 1,10)
)
```

Now join
```{r}
DF<-left_join(DF, cams, by = "Cam_SD")
```
Now let's calculate trap nights. Need to make some date vars to do so.

```{r}
DF$DTO<-mdy_hm(DF$DTO)
```

Now that we have those in as dates, let's get a summary of them:

```{r}
summary(DF$DTO)
```
Uh-oh! Some of the camera dates are incorrect, as we did not have cameras set in 2017. Let's take a look and try to fix that.

```{r}
bad_dates<-DF %>% filter(year(DTO) < 2021)
```
These are all human camera trappers. I'll try to fix, but can basically ignore if it doesn't work very well.

```{r}
good_dates<-ymd_hms(c("2021-10-17 14:50:00", "2021-10-17 14:51:00", "2021-10-04 16:23:00"))
subs<-bad_dates$subject_ids
fix<-which(DF$subject_ids %in% subs)
DF$DTO[fix]<-good_dates
summary(DF$DTO)
```
Now let's look at start and end dates
```{r}
summary(DF$start_date)
summary(DF$retrieval_date)
```
Good!

Now calculate trap nights per camera

```{r}
DF<-DF %>% mutate(
  cam_nights = as.integer(retrieval_date - start_date)
)
summary(DF$cam_nights)
```

Now get summary of cam nights per camera and phase

```{r}
CamNightSummary<-DF %>% group_by(Phase, Treatment, Cam_SD) %>% summarise(
  tot_cam_nights = max(cam_nights)
)
```

Now write some code to calculate the cumulative number of species over days for each camera.

Need a function that will:
1. Create a data frame or list to store results; that is what will be returned. Columns should include: Cam_SD, phase_treatment,      cumulative_day, cum_num_species
2. Grab the data for a single camera 
3. Put the data in order by date time of observation (DTO)
4. Assign each date to a day number for counting the number of species on each day
5. Loop through each camera and do the calculating and assigning.

Let's call it "species_cam_nights"

```{r}
species_cam_nights<-function(df){
  df<-df %>% arrange(Cam_SD, DTO)
  num_cams<-length(unique(df$Cam_SD))
  cams<-unique(df$Cam_SD)
  total_species<-length(unique(df$choice))
  df$day<-NA
  Species<-vector()
  
  newDF<-data.frame(Cam_SD = character(length(df$DTO)),
                    Phase_Treat = character(length(df$DTO)),
                    DTO = POSIXct(length(df$DTO)),
                    cumulative_day = integer(length(df$DTO)),
                    num_species = integer(length(df$DTO)))
  
  #loop to assign day, num species
  for (i in 1:num_cams){
    data<-df %>% filter(Cam_SD == cams[i])
    num_rows<-length(data$Cam_SD)
    for (a in 1:num_rows){
      
      #next 4 lines of code are wrong - need to figure out what the next empty row in the newDF data frame is and start filling from there - this code writes over the existing code. 
      newDF$Cam_SD[a]<- data$Cam_SD[a]
      newDF$Phase_Treat[a]<-data$Phase_Treat[a]
      newDF$DTO[a]<-data$DTO[a]
      Species[a]<-data$choice[a]
      if (a == 1){
        newDF$num_species[a]<- length(unique(Species))
        newDF$cumulative_day[a]<-1
      } else{
            Day_test<- (day(data$DTO[a]) == day(data$DTO[a-1]))
              if(Day_test == TRUE){
                  newDF$cumulative_day[a]<-newDF$cumulative_day[a-1]}
              else {newDF$cumulative_day[a] <- newDF$cumulative_day[a-1]+1}
            newDF$num_species[a]<-length(unique(Species[1:a]))
          }
      }
  }
  return(newDF)
}
```
Try it
```{r}
test<-species_cam_nights(DF)
```

  
  #loop to assign day
  for(i in 1:num_cams){
    data<-df %>% filter(Cam_SD == cams[i])#grab the data
    num_rows<-length(data$Cam_SD)
    data$day[1]<-1 #assign first day as 1
    start_date<-data$DTO[1]
      for(z in 2:num_rows){
        data$day[z]<-as.integer(round(difftime(data$DTO[z], data$DTO[z-1], units = "days")))
        data$cum_day[z]<-as.integer(round(difftime(data$DTO[z],start_date, units = "days")))
      }
    #now replace those subject ids in df
    sub_ids<-unique(data$subject_ids)
    swap<-which(df$subject_ids %in% sub_ids)
    df[swap,]<-data
  }
  df<-df %>% relocate(day, .after = DTO)
  return(df)
}
```

Now use the function
```{r}
DF<-species_cam_nights(DF)
```
Now write a second function to do steps 3 and 4. Call it species_per_day

```{r}
species_per_day<-function(df){
  num_cams<-length(unique(df$Cam_SD))
  cams = unique(df$Cam_SD)
  #create empty data frame
  newDF<-data.frame(Cam_SD = character(),
                    Phase_Treat = character(),
                    cumulative_day = integer(),
                    cumulative_species = integer())
  
  #loop through df by camera
  for (i in 1:num_cams){
    df_subset<-df %>% filter(Cam_SD == cams[1]) #grab data from first camera
    #add cumulative days column
    df_subset$cum_days<-cumsum(df_subset$day)
    num_rows <-length(df_subset$Cam_SD)
    for (z in 1:num_rows){
      newDF$Cam_SD[z]<-df_subset$Cam_SD[z]
      newDF$Phase_Treat[z]<-df_subset$Phase_Treat
      newDF$cumulative_day = df_subset$cum_days[z]
      if (z == 1){
        newDF$cumulative_species <-1
        pool = df_subset$choice[z]
      } else {
        pool = c(pool, df_subset$choice[z])
        newDF$cumulative_species<-length(unique(pool))
      }
      
    }
  }
  return(newDF)
}
```

